---
title: "ExploratoryDataAnalysis"
author: "Maria Williams"
date: "3/3/2022"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries:

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(ggcorrplot)
```

## Step 1) Load Dataset

This dataset was originally created by the code found in the DataPreProcessing file. To run this notebook locally, download and extract the EEGdataFull file and replace the URL below with the URL local to your machine.

```{r}
EEGdata <- readRDS("C:/Users/maria/OneDrive/Documents/BioData/BME580Group/EEGdataFull.Rda")
```

## Step 2) Visualize Sample Data

First we selected three random examples from each class, and pulled out only the first 2 seconds of each.

```{r}
#files - three control three diagnosed
these <- list("C:/Users/maria/OneDrive/Documents/BioData/BME580Group/norm/S10W1.txt", "C:/Users/maria/OneDrive/Documents/BioData/BME580Group/norm/S59LW.txt", "C:/Users/maria/OneDrive/Documents/BioData/BME580Group/norm/S155W1.txt", "C:/Users/maria/OneDrive/Documents/BioData/BME580Group/sch/022w1.txt", "C:/Users/maria/OneDrive/Documents/BioData/BME580Group/sch/642w1.txt", "C:/Users/maria/OneDrive/Documents/BioData/BME580Group/sch/585w1.txt")

#initiate sample
sample <- data.frame()

#set how many samples (128=1sec)
ok <-256

#pull in data
for (f in these){
  #read in file as list
  hey <- read.delim(f, header=FALSE)

  #reshape so each signal is a column
  mini <- matrix(unlist(hey), nrow = 7680, ncol = 16)
  mini <- as.data.frame(mini)

  #append the first chunk of values
  sample <- rbind(sample, head(mini, ok))
}

#rename columns
titles <- list("F7", "F3", "F4", "F8", "T3", "C3", "Cz", "C4", "T4", "T5", "P3", "Pz", "P4", "T6", "O1", "O2")
colnames(sample) <- titles

#add column to indicate control or diagnosed
d <-rep(0:1,each=3*ok)
sample$diagnosis <- as.factor(d)

#add a column to indicate subject
s <- rep(1:6, each=ok)
sample$subject <- as.factor(s)

#add a column to indicate obs
o <- rep(1:ok, times=6)
sample$obs <- as.factor(o)

```

Then, we visualized the waveforms to see if there were any visible trends or differences between the classes.

```{r}

P3plot = ggplot(sample, aes(x=obs, y=P3, group=subject, color=diagnosis)) +
  geom_line(show.legend = FALSE)+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), plot.caption = element_text(hjust = 0.5, face= "italic"))+
  scale_color_discrete(name = 'Diagnosis', labels=c("non-symptomatic", "schizophrenic" ))+
  labs(title="P3 Region", y="voltage")

O2plot = ggplot(sample, aes(x=obs, y=O2, group=subject, color=diagnosis)) +
  geom_line()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), plot.caption = element_text(hjust = 0.5, face= "italic"), legend.position="bottom")+
  scale_color_discrete(name = 'Diagnosis:', labels=c("control", "schizophrenic" ))+
  labs(title="O2 Region", y="voltage")

grid.arrange(P3plot, O2plot, heights=c(1, 1.6))
```


This quick visualization does not show any obvious differences in the classes in amplitude or frequency, or any noticable trends.

## Step 3) Correlation Plots

We also created some correlation plots of the six random samples to see if we could visually identify any patterns of correlation.

```{r}
plots <- list()
d="Control"

for (i in 1:6){
  #make title
  if (i>3){d="Schizophrenic"}
  yay <- cat("Subject", i, "(", d, ")")
  
  #pull out one subject
  this <- sample[sample$subject==i,]
  this <- this[1:16]

  #get the correlations
  corr <- round(cor(this), 1)
  
  #plot
  plots[[i]] <- ggcorrplot(corr, show.legend = FALSE) +
    labs(title = yay)
}

grid.arrange(plots[[1]], plots[[2]])


```



## Step 4) Full Data Box Plots

To explore the full dataset, we first gather the summary data of each sensor for each subject.

```{r}
#summarize across the entire brain activity for each subject
allsum <- summary(t(EEGdata))

#create as dataframe
allsum <- as.data.frame.matrix(allsum)
#rename columns since they have a strange leading character
for (i in 1:84){
    names(allsum)[i]<- paste("subject", i, sep="_")
}
#strip out numeric values
for (i in 1:84){
  allsum[,i]<- as.numeric(gsub(".*:", "", allsum[,i]))
}

#transpose so subjects are rows and summary stats are columns
allsum <- t(allsum)
colnames(allsum)<- c("Min", '1stQ', 'Median', 'Mean', '3rdQ', 'Max')
allsum <- as.data.frame.matrix(allsum)

#add column for range
allsum$Range <- allsum$Max-allsum$Min

#add column to indicate control or diagnosed
d <-rep(c(0,1),times=c(39,45))
allsum$diagnosis <- as.factor(d)

```

We then graph the resulting table.

```{r}

Minplot <- ggplot(allsum, aes(x=Min, group=diagnosis, fill=diagnosis)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title="Lowest Recorded Voltage", x=NULL) +
  theme(text = element_text(size=20), axis.text.y=element_blank(), axis.title.y=element_blank())
Maxplot <- ggplot(allsum, aes(x=Max, group=diagnosis, fill=diagnosis)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title="Highest Recorded Voltage", x=NULL) +
  theme(text = element_text(size=20), axis.text.y=element_blank(), axis.title.y=element_blank())
Meanplot <- ggplot(allsum, aes(x=Mean, group=diagnosis, fill=diagnosis)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title="Average Recorded Voltage", x=NULL) +
  theme(text = element_text(size=20), axis.text.y=element_blank(), axis.title.y=element_blank())
Rangeplot <- ggplot(allsum, aes(x=Range, group=diagnosis, fill=diagnosis)) +
  geom_boxplot() +
  labs(title="Range of Recorded Voltage", x=NULL) +
  theme(text = element_text(size=20), axis.text.y=element_blank(), axis.title.y=element_blank(), legend.position="bottom")+
  scale_fill_discrete(name="Diagnosis", labels = c("control", "schizophrenic"))

grid.arrange(Minplot, Maxplot, Meanplot, Rangeplot, nrow=4, heights=c(1, 1, 1, 1.3))
                 
```
