---
title: "FeatureEngineering"
author: "Maria Williams"
date: "3/23/2022"
output: html_document
---

Libraries:

```{r}
library(readxl)
library(dbplyr)
library(ggplot2)
```

## Step 1) Load Dataset

This dataset was originally created by the code found in the DataPreProcessing file. To run this notebook locally, download and extract the EEGdataFull file and replace the URL below with the URL local to your machine.

```{r}
EEGdata <- readRDS("C:/Users/maria/OneDrive/Documents/BioData/BME580Group/EEGdataFull.Rda")
```



## Step 2) Feature Engineering

For each of the sixteen channels, we will extract:

* the average voltage 
* the minimum voltage
* the maximum voltage
* the range of voltage measurements
* the number of peaks in the voltage

### Custom Function "MyWave"

This function currently returns the number of peaks from a row, but hopefully we can extract some more data to return as well.

```{r}
#note: misses 'peaks' that sustain
MyWave <- function (wiggle){
  wigglelength <- length(wiggle)
  holder <- vector(mode = "list", length = wigglelength)

  #check first value
  if (wiggle[1]>wiggle[2]){
    holder[1]=="p"
  }

  #check last value
  if (wiggle[wigglelength]>wiggle[wigglelength-1]){
    holder[wigglelength]=="p"
  }

  #check inbetween values - doesn't work if start at 2
  for (i in 3:wigglelength-1){
    #cat(wiggle[i-1], wiggle[i], wiggle[i+1], '\n')
    if (wiggle[i]>wiggle[i+1] && wiggle[i]>wiggle[i-1]){
      holder[i]='p'
    }
  }

  return (sum(holder=='p'))
}

```


Here is a block of code to test it

```{r}
#testing - pulling out .5 second from three subjects
hey <- EEGdata[1:3,1:64]
#transpose (switch rows and columns)
mini <- as.data.frame(t(hey))
#quick visualization
plot(mini[,2], type = 'l', xlab="", ylab="")
#switch back to call function
mini <- hey

#function call
mini$peaks <- apply(mini, 1, MyWave)
print(mini$peaks)

```



### Feature Engineering on full dataset

```{r}
#start new dataframe with labels
EEGnew <- data.frame(EEGdata[,122881])
names(EEGnew)[1]<- "diagnosis"

#list of sensors
sensors <- list("F7", "F3", "F4", "F8", "T3", "C3", "Cz", "C4", "T4", "T5", "P3", "Pz", "P4", "T6", "O1", "O2")
#list of new features
columns <- list("avg", "min", "max", "range", "numpeaks")
#counter for new columns (column 1 is the label)
colcount = 2
#counter for subsets
chunkcount = 0

for (s in sensors){
  #print(names(EEGdata)[chunkcount+1])
  #print(names(EEGdata)[chunkcount+7680])
  
  #subset the data from a single sensor
  chunk <- EEGdata[(chunkcount+1):(chunkcount+7680)]
  #get stats
  EEGnew$avg <- apply(chunk, 1, mean)
  EEGnew$min <- apply(chunk, 1, min)
  EEGnew$max <- apply(chunk, 1, max)
  EEGnew$range <- EEGnew$max - EEGnew$min
  #get waveform details
  EEGnew$peaks <- apply(chunk, 1, MyWave)
  
  #increment subset
  chunkcount<-chunkcount+7680
  
  #rename columns
  for (c in columns){
    names(EEGnew)[colcount]<- paste(s, c, sep="_")
    colcount <- colcount+1
  }
}

```

Separate chunk for correlation metrics

```{r}
#create the column names for the correlations
#loops through the sensor names but omits 1:1 combos and flipped duplicates
titles2= list()
for (m in titles){
  for (w in titles){
    if (m!=w){
      t=paste(w, m, "Cor", sep='_')
      flip=paste(m, w, "Cor", sep='_')
      if (flip %in% titles2){}
      else {
        titles2[[length(titles2) + 1]]=t
      }
    }
  }
}

#instantiate the correlation dataframe separately
corrTab <-data.frame(matrix(ncol=0, nrow=120))

#for each subject, get the correlation values of its sensors
for (i in 1:84){
  #pull out one subject (remove dependent variable)
  this <- EEGdata[i,1:122880]
  #arrange it by sensor
  this <- matrix(unlist(this), nrow = 7680, ncol = 16)
  #get the correlations
  corr <- round(cor(this), 2)
  corr <- corr[lower.tri(corr, diag=FALSE)]
  #zero out values <.5
  corr[corr<.5]=0
  #add to dataframe
  corrTab[,i] = corr
}

#reshape so observations are rows
corrTab=as.data.frame(t(corrTab))
#rename columns
names(corrTab)=titles2
#add to dataset
EEGnew<-cbind(EEGnew, corrTab)

```


###Export File

Final File includes 84 observations with 201 factors:
* diagnosis flag
* 80 sensor stats (5 stats for each of the 16 sensors)
* 120 correlation values (for each pair of 16 sensors)

```{r}
#export file as rda
#saveRDS(EEGnew,"C:/Users/maria/OneDrive/Documents/BioData/SchizophreniaEEG/EEGnew.rds")
```

