---
title: "FeatureEngineering"
author: "Maria Williams"
date: "3/23/2022"
output: html_document
---

Libraries:

```{r}
library(readxl)
library(dbplyr)
library(ggplot2)
```

## Step 1) Load Dataset

This dataset was originally created by the code found in the DataPreProcessing file. To run this notebook locally, download and extract the EEGdataFull file and replace the URL below with the URL local to your machine.

```{r}
EEGdata <- readRDS("C:/Users/maria/OneDrive/Documents/BioData/BME580Group/EEGdataFull.Rda")
```



## Step 2) Feature Engineering

### Custom Function "MyWave"

This function receives one waveform and returns:
* the number of valleys in the wave
* the number of peaks in the wave
* the average voltage of the valleys of the wave
* the average voltage of the peaks of the wave
* the range of voltage of the valleys of the wave
* the range of voltage of the peaks of the wave

```{r}
MyWave <- function (wiggle){
  wigglelength <- length(wiggle)
  peaks <- rep(NA, wigglelength)
  vals <- rep(NA, wigglelength)

  #setting index value
  #using while loop instead of for loop because we will skip indexes
  i=1
  while (i<=wigglelength){
    
    #first set the position to compare
    #usually the next value but sometimes values are sustained
    n=1
    #print(cat(i, wigglelength, wiggle[i], wiggle[i+n], n))
    while (i<wigglelength && wiggle[i]==wiggle[i+n] && i+n<=wigglelength){
      n=n+1
    }

    #if its the first one
    if (i==1){
      if (wiggle[i]<wiggle[i+n]){
        vals[i]=wiggle[i]
      }
      else if (wiggle[i]>wiggle[i+n]){
        peaks[i]=wiggle[i]
      }
      else{
        #should not happen
        print("First value not evaluated")
      }
    }
    
    #if its in the middle
    else if (i>1 && i<wigglelength && i+n<=wigglelength){
      if (wiggle[i-1] > wiggle[i] && wiggle[i] < wiggle[i+n]){
        vals[i]=wiggle[i]
      }
      else if (wiggle[i-1] < wiggle[i] && wiggle[i] > wiggle[i+n]){
        peaks[i]=wiggle[i]
      }
      else{
        #should be just in the middle of ascent/descent
      }
    }
    
    #if its the last
    else if (i==wigglelength || i+n>=wigglelength){
      if (wiggle[i]<wiggle[i-1]){
        vals[i]=wiggle[i]
      }
      else if (wiggle[i]>wiggle[i-1]){
        peaks[i]=wiggle[i]
      }
      else{
        #should only happen if the last value is sustained
        print(cat("current value:", wiggle[i], "\nComparison value:", wiggle[i-1]))
      }
    } 
    
    #increment to next non-equal index
    i=i+n
  }
  #calculate number of peaks/valleys
  numv <- wigglelength - sum(is.na(vals))
  nump <- wigglelength - sum(is.na(peaks))
  #calculate average peak/valley
  avgv <- mean(vals, na.rm=TRUE)
  avgp <- mean(peaks, na.rm=TRUE)
  #calculate range of peaks/valleys
  rangev <- max(vals, na.rm=TRUE)-min(vals, na.rm=TRUE)
  rangep <- max(peaks, na.rm=TRUE)-min(peaks, na.rm=TRUE)
  #calculate average / range of wavelength
  #calculate avg/range of descents/ascents
  #avgd
  #avga
  #ranged
  #rangea
  
  
  return (c(numv, nump, avgv, avgp, rangev, rangep))
}
```

Here is a block of code to test it

```{r}
#testing - pulling out .5 second from three subjects
hey <- EEGdata[1:3,1:64]
#transpose (switch rows and columns)
mini <- as.data.frame(t(hey))
#quick visualization
plot(mini[,1], type = 'l', xlab="", ylab="")
#switch back to call function
mini <- hey

#function call
miniwave <- apply(mini, 1, MyWave)
#mini$peaks <- apply(mini, 1, MyWave)
#reshape so observations are rows
miniwave<-as.data.frame(t(miniwave))
mini<-cbind(mini, miniwave)

```



### Feature Engineering on full dataset

For each of the sixteen channels, we will extract:

* the average voltage 
* the minimum voltage
* the maximum voltage
* the range of voltage measurements

We will use the MyWave function to extract:

* the number of valleys in the wave
* the number of peaks in the wave
* the average voltage of the valleys of the wave
* the average voltage of the peaks of the wave
* the range of voltage of the valleys of the wave
* the range of voltage of the peaks of the wave

```{r}
#start new dataframe with labels
EEGnew <- data.frame(EEGdata[,122881])
names(EEGnew)[1]<- "diagnosis"

#list of sensors
sensors <- list("F7", "F3", "F4", "F8", "T3", "C3", "Cz", "C4", "T4", "T5", "P3", "Pz", "P4", "T6", "O1", "O2")
#list of new features
columns <- list("avg", "min", "max", "range", "valnum", "peaknum","valavg", "peakavg", "valrange", "peakrange")
#counter for new columns (column 1 is the label)
colcount = 2
#counter for subsets
chunkcount = 0

for (s in sensors){
  #print(names(EEGdata)[chunkcount+1])
  #print(names(EEGdata)[chunkcount+7680])
  
  #subset the data from a single sensor
  chunk <- EEGdata[(chunkcount+1):(chunkcount+7680)]
  #get stats
  EEGnew$avg <- apply(chunk, 1, mean)
  EEGnew$min <- apply(chunk, 1, min)
  EEGnew$max <- apply(chunk, 1, max)
  EEGnew$range <- EEGnew$max - EEGnew$min
  #get waveform details
  EEGwave <- apply(chunk, 1, MyWave)
  EEGwave <- as.data.frame(t(EEGwave))
  #add to dataset
  EEGnew <- cbind(EEGnew, EEGwave)
  
  #increment subset
  chunkcount<-chunkcount+7680
  
  #rename columns
  for (c in columns){
    names(EEGnew)[colcount]<- paste(s, c, sep="_")
    colcount <- colcount+1
  }
}

```

Separate chunk for correlation metrics

```{r}
#create the column names for the correlations
#loops through the sensor names but omits 1:1 combos and flipped duplicates
titles2= list()
for (m in titles){
  for (w in titles){
    if (m!=w){
      t=paste(w, m, "Cor", sep='_')
      flip=paste(m, w, "Cor", sep='_')
      if (flip %in% titles2){}
      else {
        titles2[[length(titles2) + 1]]=t
      }
    }
  }
}

#instantiate the correlation dataframe separately
corrTab <-data.frame(matrix(ncol=0, nrow=120))

#for each subject, get the correlation values of its sensors
for (i in 1:84){
  #pull out one subject (remove dependent variable)
  this <- EEGdata[i,1:122880]
  #arrange it by sensor
  this <- matrix(unlist(this), nrow = 7680, ncol = 16)
  #get the correlations
  corr <- round(cor(this), 2)
  corr <- corr[lower.tri(corr, diag=FALSE)]
  #zero out values <.5
  corr[corr<.5]=0
  #add to dataframe
  corrTab[,i] = corr
}

#reshape so observations are rows
corrTab=as.data.frame(t(corrTab))
#rename columns
names(corrTab)=titles2
#add to dataset
EEGnew<-cbind(EEGnew, corrTab)

```


###Export File

Final File includes 84 observations with 281 factors:
* diagnosis flag
* 64 sensor stats (4 stats for each of the 16 sensors)
* 96 waveform stats (6 stats for each of the 16 sensors)
* 120 correlation values (for each pair of 16 sensors)

```{r}
#export file as rda
#saveRDS(EEGnew,"C:/Users/maria/OneDrive/Documents/BioData/SchizophreniaEEG/EEGnew.rds")
```

